---
title: "Coursera reproduceable research week 2 assignment"
author: "Edgar Ekure"
date: "March 12, 2017"
output: html_document
---

##First we load R packages to be used


```{r echo=TRUE}
library(scales)
library(Hmisc)
library(ggplot2)
```


##Read the csv file "activity"


```{r}
data_act <- read.csv('activity.csv')
```


##What is mean total number of steps taken per day?


```{r}
totaldailysteps <-  tapply(data_act$steps, data_act$date, sum, na.rm=TRUE)
```


###Make a histogram of the total number of steps taken each day


```{r}
qplot(totaldailysteps, xlab='Total steps per day', ylab='Frequency using binwith 500', binwidth=500)
```


###Calculate and report the mean and median of the total number of steps taken per day


```{r echo=TRUE}
 meantotaldailysteps <- mean(totaldailysteps, na.rm = TRUE)
meantotaldailysteps
mediantotaldailysteps <- median(totaldailysteps, na.rm = TRUE)
mediantotaldailysteps
```


##What is the average daily activity pattern?


### 1.  Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)


```{r}
averagenumsteps <- aggregate(x=list(ave_steps=data_act$steps), by=list(interval=data_act$interval), FUN=mean, na.rm=TRUE)
ggplot(data=averagenumsteps, aes(x=interval, y=ave_steps)) +
    geom_line() +
    xlab("5-minute interval") +
    ylab("average number of steps taken") 
```


### 2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?


```{r}
maxsteps <- which.max(averagenumsteps$ave_steps)
maxsteps
maxstepstime <-  gsub("([0-9]{1,2})([0-9]{2})", "\\1:\\2", averagenumsteps[maxsteps,'interval'])
maxstepstime
```


##Imputing missing values


### 1.Number of missing values


```{r}
numofnas <- length(which(is.na(data_act$steps)))
numofnas
```


### 2. Devise a strategy for filling in all of the missing values in the dataset.


### Create a new dataset that is equal to the original dataset but with the missing data filled in.


```{r}
imputedactdata <- data_act
imputedactdata$steps <- impute(data_act$steps, fun=mean)
```


## Make a histogram of the total number of steps taken each day


```{r}
imputedtotaldailysteps <- tapply(imputedactdata$steps, imputedactdata$date, sum)
qplot(imputedtotaldailysteps, xlab='Total steps per day (Imputed)', ylab='Frequency using binwith 500', binwidth=500)
```


####Calculate and report the mean and median total number of steps taken per day.


```{r}
meanimputedtotaldailysteps <- mean(imputedtotaldailysteps)
meanimputedtotaldailysteps
medianimputedtotaldailysteps <- median(imputedtotaldailysteps)
medianimputedtotaldailysteps
```


##Are there differences in activity patterns between weekdays and weekends?


```{r}
imputedactdata$dateType <-  ifelse(as.POSIXlt(imputedactdata$date)$wday %in% c(0,6), 'weekend', 'weekday')
```


###Make a panel plot containing a time series plot


```{r}
meanimputedact_data <- aggregate(steps ~ interval + dateType, data=imputedactdata, mean)
ggplot(meanimputedact_data, aes(interval, steps)) + 
    geom_line() + 
    facet_grid(dateType ~ .) +
    xlab("5-minute interval") + 
    ylab("avarage number of steps")
```


